Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-29 ms.tex

\begin{abstract} 
  Social learning %essential to human survival---it 
  is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four distinct and theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of social learning in populations with cognitively-realistic individual-learning
  capacities. % We used evolutionary agent-based modeling to track the evolution of social learning versus purely individual learning.
  %how often social, rather than individual,   learning evolved. 
  Agents could perform one of several behaviors, modeled as multi-armed bandit, to acquire payoffs.
  Individual learning used the softmax algorithm.  Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning.  We found that different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning, whereas larger decision-set size promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This study shows the importance of clearly
  operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
\footnote{This document contains
\total{citnum} references.}  
\end{abstract}

>>>>

\begin{abstract} 
  Social learning is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four distinct and theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of  social learning in populations with cognitively realistic individual learning
  capacities.  
  Agents could perform one of several behaviors, modeled as multi-armed bandit, to acquire payoffs.
  Individual learning used the softmax algorithm.  Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning.  We found that different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning, whereas larger decision-set size promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This study shows the importance of clearly
  operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
% \footnote{This document contains
% \total{citnum} references.}  
\end{abstract}
